# Manual SCE classification results

> Compile with `Rscript -e 'rmarkdown::render("meta_analysis.Rmd")'`

```{r, echo = F, warning = F, message = F}
library(data.table)
source("/srv/shiny-server/kinanen/SCE-Detection/global.R")
```

Read in results:

```{r}
sce <- rbind(cbind(author = "Ashley", fread("/srv/shiny-server/kinanen/SCE-Detection/Results/Ashley_results.txt")),
             cbind(author = "Sascha", fread("/srv/shiny-server/kinanen/SCE-Detection/Results/Sascha_results.txt")),
             cbind(author = "Karen",  fread("/srv/shiny-server/kinanen/SCE-Detection/Results/Karen_results.txt")),
             cbind(author = "Venla",  fread("/srv/shiny-server/kinanen/SCE-Detection/Results/Venla_results.txt")))
```

This is a list of all SCEs found:

```{r}
all_sce <- sce[, .(num_authors = length(unique(author))), by = .(sample, cell, chrom, end)]
```


## General stats

```{r progress_by_chrom, fig.width=6, fig.height=3}
ggplot(sce) + aes(chrom, fill = author) + geom_bar() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This is how many people agreed on SCEs per chromosome:

```{r num_of_sce_supporters_by_chrom, fig.width = 8, fig.height = 6}
ggplot(all_sce) + aes(num_authors) + geom_histogram(binwidth = 1) + facet_wrap(~chrom)
```

At last we can derive a confusion table of how often authors agree:

```{r}
conftab = NULL
for (author1 in unique(sce$author)) {
  for (author2 in unique(sce$author)) {
    x = sce[chrom %in% paste0("chr", 1:5) & author %in% c(author1, author2), length(author), 
                 by = .(cell_name, chrom, end)]
    if (author1 == author2) {
      n = nrow(x)
    } else {
      n = nrow(x[V1>1,])
    }
    conftab = rbind(conftab,
                    data.table(author1 = author1, 
                               author2 = author2,
                               n = n))
  }
}
conftab <- merge(conftab, conftab[, .(N = max(n)), by = author1], by = "author1")
conftab <- conftab[, r:= n/N][]
ggplot(conftab) + aes(author1, author2, fill = r) + geom_raster() + scale_fill_gradientn(colours = c("red1","white","dodgerblue"))
```

# Inspecting problematic SCEs

To make plots, let's read the raw Mosaicatcher data

```{r}
counts <- fread("/srv/shiny-server/kinanen/SCE-Detection/children.500kb.txt", showProgress = F)
counts <- initialize_data(counts)
```

This is a function to plot a bunch of cells where authors don't agree on an SCE. 
It shows the same chromosome of up to 8 cells which contain such an SCE and also 
labels the potential SCE by the number of authors that clicked it.

```{r}
inspect_problematic_sces <- function(CHR) {
  selection_cells  = unique(all_sce[chrom == CHR & num_authors < 4,.(sample, cell)])[1:8]
  selection_sce     = sce[selection_cells, on = .(sample,cell)][chrom == CHR]
  selection_counts  = counts[selection_cells, on = .(sample,cell)][chrom == CHR]
  selection_all_sce = all_sce[selection_cells, on = .(sample,cell)][chrom == CHR]
  main_plot(selection_counts) + 
    geom_vline(data = selection_sce,  aes(xintercept = end)) +
    geom_label(data = selection_all_sce, aes(label = num_authors, x = end, y = Inf, hjust = 1, vjust = 1))
}
```

```{r problematic_sces_chr1, fig.width = 10, fig.height = 8}
inspect_problematic_sces("chr1")
```

```{r problematic_sces_chr2, fig.width = 10, fig.height = 8}
inspect_problematic_sces("chr2")
```

```{r problematic_sces_chr3, fig.width = 10, fig.height = 8}
inspect_problematic_sces("chr3")
```

```{r problematic_sces_chr4, fig.width = 10, fig.height = 8}
inspect_problematic_sces("chr4")
```




END.